---
title: "Exploring Patient Table"
output: rmarkdown::github_document
---
Author: [Mahdi Moqri](https://www.moqri.com/)

The patinet table is a core part of the eICU-CRD and contains all information related to tracking patient unit stays. The table also contains patient demographics and hospital level information.


### Loading libraries
```{r message=FALSE}
library(data.table)
library(dplyr)
library(magrittr)
#library(plyr)
library(knitr)
library(ggplot2)
library(visreg)
library(rpart)
```

### Examining the patient table
The uniquePid column identifies a single patient across multiple stays. Let's look at a single uniquepid.

```{r}
# data_folder is the local path to your data
df=fread('patient.csv')
cols=c('uniquepid', 'gender', 'age', 'ethnicity','hospitaldischargestatus')
kable(head(df %>% select(cols)))
```
### Age
```{r message=FALSE, warning=FALSE}
df %<>% mutate(age_impute= (ifelse(age=='> 89',90, as.numeric(age))))
age=df[!is.na(df$age_impute),]$age_impute
hist(age,100)
median(age)
data.frame(table(df$age))
```

### Gender
```{r}
gender=table(df$gender)[c('Female','Male')]
gender
pie(gender)

```

### Ethnicity
```{r, catche=TRUE}
df_ethnicity =df %>% subset(ethnicity!='')
ethnicity=df_ethnicity$ethnicity
kable(table(ethnicity))
barchart(table(ethnicity))
```

### Status at discharge
```{r}
unit_discharge=df$unitdischargestatus
table(unit_discharge)
pie(table(unit_discharge))
hostpial_discharge=df$hospitaldischargestatus
kable(table(hostpial_discharge))
pie(table(hostpial_discharge))
```

###relationship between variables
```{r}
df<-na.omit(df)
my_data<- df %>% select (-c(age))
CTree <- rpart(hospitaldischargestatus~., data=my_data, method="class")
rpart.plot(CTree, tweak = 1.5, cex=.6, type=5,extra = 106)

```


### Age vs Outcome
```{r message=FALSE, warning=FALSE}

#ggplot(data = df) +
 # geom_bar(mapping = aes(x = age, y = ..prop.., group = 1), stat = "count")
#Hospital_dischargestatus=df$hospitaldischargestatus
#ggplot(df, aes(x = age)) +  
#  geom_bar(aes(y = (..count..)/sum(..count..)))
#ggplot(df, aes(fill=Hospital_dischargestatus, y=..prop.., x=age)) + 
  #  geom_bar(position="dodge", stat="identity")
#count(df$hospitaldischargestatus %in% c("Expired"))


df = df %>% mutate(hospitaldischargestatus = ifelse(hospitaldischargestatus  %in% c("Expired"), 1, 0))
cps85_glm <- glm(hospitaldischargestatus ~  age , 
                 family="binomial", 
                 data=df)
visreg(cps85_glm, "age", 
       gg = TRUE, 
       scale="response")
```

```{r}
require(dplyr)

NewTable=df %>% group_by(age_impute,gender) %>%
  summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100) 
NewTable<-na.omit(NewTable)
NewTable<-filter (NewTable, gender %in% c("Male","Female"))
plot(NewTable$`age_impute`, NewTable$percentage, main = "Age vs Death",
     xlab = "Age", ylab = "Death probability",
     pch = 19, frame = FALSE)

```
```{r}
ggplot(NewTable, aes(x=age_impute, y=percentage, group=gender)) +
  geom_point(aes(color=gender))
```



###Age-Gender-Outcome
```{r}
b <- c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
names=c("0-5","5-10","10-15","15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
NewTable$BreakedAge=cut(NewTable$age_impute, breaks = b, labels = names)
NewTable<-na.omit(NewTable)
SecondTable=NewTable %>% group_by(BreakedAge,gender) %>%
  summarise(expiredTotal=sum(percentage)) 
ggplot(SecondTable, aes(x=BreakedAge, y=expiredTotal/5, group=gender)) +
  geom_line(aes(color=gender))+
  geom_point(aes(color=gender))
```


###Race-Outcome
```{r}
ethnicityTable=df %>% group_by(age_impute,ethnicity,gender) %>%
  summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100) 
ethnicityTable<-filter (ethnicityTable, gender %in% c("Male","Female"))
ethnicityTable<-na.omit(ethnicityTable)
ggplot(data = ethnicityTable) +
  geom_boxplot(mapping = aes(x = ethnicity, y = percentage)) +
  coord_flip()

```
###Age-race-Outcome
```{r}

ethnicityTable<-filter (ethnicityTable, ethnicity %in% c("Caucasian","African American"))
plot(NewTable$`age_impute`, NewTable$percentage, main = "Race vs Death",
     xlab = "Age", ylab = "Death probability",
     pch = 19, frame = FALSE)

b <- c(15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
names=c("15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
ethnicityTable$BreakedAge=cut(ethnicityTable$age_impute, breaks = b, labels = names)
ethnicityTable<-na.omit(ethnicityTable)
ethnicitySecondTable=ethnicityTable %>% group_by(BreakedAge,ethnicity) %>%
  summarise(expiredTotal=sum(percentage)) 
ggplot(ethnicitySecondTable, aes(x=BreakedAge, y=expiredTotal/5, group=ethnicity)) +
  geom_line(aes(color=ethnicity))+
  geom_point(aes(color=ethnicity))
ggplot(ethnicityTable, aes(x=age_impute, y=percentage, group=ethnicity)) +
  geom_point(aes(color=ethnicity))


ggplot(data = ethnicityTable) +
  geom_boxplot(mapping = aes(x = ethnicity, y = percentage)) +
  coord_flip()


ggplot(data = ethnicityTable, aes(x = ethnicity, y = percentage)) + geom_boxplot(aes(fill=gender))
```
###T-TEST
```{r}
dfEthOut=df%>%select(c(ethnicity,hospitaldischargestatus))
dfEthOut <-na.omit(dfEthOut)
dfEthOut=dfEthOut%>%mutate(output=ifelse(hospitaldischargestatus  %in% c("Expired"), 1, 0))
White=dfEthOut%>%subset(ethnicity=="Caucasian")
Black=dfEthOut%>%subset(ethnicity=="African American")

t.test(White$output,Black$output)
```

###disease-outcome
```{r}
DiseaseTable=df %>% group_by(apacheadmissiondx) %>%
  summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100) 
DiseaseTable<-na.omit(DiseaseTable)
DiseaseTable<-filter (DiseaseTable, total>50)
DiseaseTable<-filter (DiseaseTable, percentage>20)


DiseaseTable$diseaseId=as.numeric(as.factor(DiseaseTable$apacheadmissiondx))

ggplot(data=DiseaseTable,aes(x=reorder(diseaseId,percentage),y=percentage)) + 
  geom_bar(stat ='identity',aes(fill=percentage))+
  coord_flip() + 
  theme_grey() + 
  scale_fill_gradient(name="percentage")+
  labs(title = 'disease vs percentage',
       y='percentage',x='disease')+ 
  geom_hline(yintercept = mean(DiseaseTable$percentage),size = 1, color = 'blue')
```
```{r}
DiseaseTable=df %>% group_by(apacheadmissiondx) %>%
  summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100) 
DiseaseTable<-na.omit(DiseaseTable)
# top_fifty <- DiseaseTable %>%
#          filter(rank(desc(total))<=40)%>%
#          filter(percentage>10)
#         
#DiseaseTable=head(sort(DiseaseTable$total,decreasing=TRUE), n = 50)
#DiseaseTable <- DiseaseTable[order(-DiseaseTable$total) , ]

#DiseaseTable=DiseaseTable%>%top_n(50)
#DiseaseTable=DiseaseTable%>%filter (percentage>20)
DiseaseTable<-filter (DiseaseTable, total>1000)
DiseaseTable<-filter (DiseaseTable, percentage>15)
# 
# 
DiseaseTable$diseaseId=as.numeric(as.factor(DiseaseTable$apacheadmissiondx))

ggplot(data=DiseaseTable,aes(x=reorder(diseaseId,percentage),y=percentage)) +
  geom_bar(stat ='identity',aes(fill=percentage))+
  coord_flip() +
  theme_grey() +
  scale_fill_gradient(name="percentage")+
  labs(title = 'disease vs percentage',
       y='percentage',x='disease')+
 geom_hline(yintercept = mean(DiseaseTable$percentage),size = 1, color = 'blue')
```
