---
title: "Exploring Patient Table"
output: rmarkdown::github_document
---
Author: [Mahdi Moqri](https://www.moqri.com/)

The patinet table is a core part of the eICU-CRD and contains all information related to tracking patient unit stays. The table also contains patient demographics and hospital level information.


### Loading libraries
```{r message=FALSE}
library(data.table)
library(dplyr)
library(magrittr)
#library(plyr)
library(knitr)
library(ggplot2)
library(visreg)
library(rpart)
library("caret")
library(rpart.plot)
```

### Examining the patient table
The uniquePid column identifies a single patient across multiple stays. Let's look at a single uniquepid.

```{r}
# data_folder is the local path to your data
dfPatient=fread('patient.csv')
#dfHospital=fread('hospital.csv')
dfapachePatientResult=fread('apachePatientResult.csv')
#df= merge(x = dfPatient, y = dfHospital, by = "hospitalid", all = TRUE)
df= merge(x = dfPatient, y = dfapachePatientResult, by = "patientunitstayid", all = TRUE)

```
### Age
```{r message=FALSE, warning=FALSE}
df %<>% mutate(age_impute= (ifelse(age=='> 89',90, as.numeric(age))))
age=df[!is.na(df$age_impute),]$age_impute
```

###Logistic regression
```{r}

NewData=df%>%select(c(age_impute,ethnicity,apachescore))%>%subset(ethnicity %in% c("Caucasian","African American"))%>%subset(apachescore!='')%>%group_by(ethnicity, age_impute) %>% summarise(meanApachi=mean(apachescore))
NewData<-na.omit(NewData)
ggplot(NewData, aes(x=age_impute, y=meanApachi)) +
 geom_point(aes(color=ethnicity))


########5years
#b <- c(15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
#names=c("15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
b <- c(10,20,30,40,50,60,70,80,90)
names=c("10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90")
NewData$BreakedAge=cut(NewData$age_impute, breaks = b, labels = names)
NewData<-na.omit(NewData)



ggplot(NewData,aes(x=BreakedAge, y=meanApachi,width=.45,fill=factor(ethnicity)))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_discrete(name="ethnicity") +
  xlab("Age")+ylab("Apachi Score")






```

