---
title: "Exploring Patient Table"
output: rmarkdown::github_document
---
Author: [Mahdi Moqri](https://www.moqri.com/)

The patinet table is a core part of the eICU-CRD and contains all information related to tracking patient unit stays. The table also contains patient demographics and hospital level information.


### Loading libraries
```{r message=FALSE}
library(data.table)
library(dplyr)
library(magrittr)
#library(plyr)
library(knitr)
library(ggplot2)
library(visreg)
library(rpart)
```

### Examining the patient table
The uniquePid column identifies a single patient across multiple stays. Let's look at a single uniquepid.

```{r}
# data_folder is the local path to your data
df=fread('patient.csv')
cols=c('uniquepid', 'gender', 'age', 'ethnicity','hospitaldischargestatus')
kable(head(df %>% select(cols)))
```
### Age
```{r message=FALSE, warning=FALSE}
df %<>% mutate(age_impute= (ifelse(age=='> 89',90, as.numeric(age))))
age=df[!is.na(df$age_impute),]$age_impute
hist(age)
hist(age)
median(age)
data.frame(table(df$age))
df_hist =df %>%subset(hospitaldischargestatus %in% c("Expired","Alive") )
ggplot(df_hist, aes(x=age, fill=hospitaldischargestatus)) +
geom_histogram(stat="count")     
#b <- c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
#names=c("0-5","5-10","10-15","15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
b <- c(10,20,30,40,50,60,70,80,90)
names=c("10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90")
df_hist$BreakedAge=cut(df_hist$age_impute, breaks = b, labels = names)
df_hist<-na.omit(df_hist)
SecondHistTable=df_hist %>% group_by(BreakedAge) 
ggplot(SecondHistTable, aes(x=BreakedAge, fill=hospitaldischargestatus)) +
geom_histogram(stat="count")  +
  xlab("Age")+ylab("Freq")
```

### Gender
```{r}
gender=table(df$gender)[c('Female','Male')]
gender
pie(gender)


SecondGenderTable=df%>%select (gender, hospitaldischargestatus)%>%subset(gender %in% c('Female','Male'))%>%subset(hospitaldischargestatus %in% c("Expired","Alive") ) %>% group_by(gender) 
SecondGenderTable=na.omit(SecondGenderTable)
ggplot(SecondGenderTable, aes(x=gender, fill=hospitaldischargestatus)) +
geom_histogram(stat="count")  +
  xlab("Gender")+ylab("Freq")



```

### Ethnicity
```{r, catche=TRUE}
df_ethnicity =df %>% subset(ethnicity!='')
ethnicity=df_ethnicity$ethnicity
kable(table(ethnicity))
barplot(table(ethnicity))
```

### Status at discharge
```{r}
unit_discharge=df$unitdischargestatus
table(unit_discharge)
pie(table(unit_discharge))
hostpial_discharge=df$hospitaldischargestatus
kable(table(hostpial_discharge))
pie(table(hostpial_discharge))
```

###relationship between variables
```{r}
# df<-na.omit(df)
# my_data<- df %>% select (-c(age))
# CTree <- rpart(hospitaldischargestatus~., data=my_data, method="class")
# rpart.plot(CTree, tweak = 1.5, cex=.6, type=5,extra = 106)

```


### Age vs Outcome
```{r message=FALSE, warning=FALSE}


df = df %>% mutate(hospitaldischargestatusId = ifelse(hospitaldischargestatus  %in% c("Expired"), 0, 1))
dfNewTable=df%>%select(c("hospitaldischargestatusId","gender","age_impute"))
NewTableAge=dfNewTable %>% group_by(age_impute) %>%
  summarise(expired=sum(hospitaldischargestatusId=="0"),total=sum(hospitaldischargestatusId %in%c("0","1")),percentage=expired/total*100) 
NewTableAge<-na.omit(NewTableAge)
plot(NewTableAge$`age_impute`, NewTableAge$percentage, main = "Age vs Death",
     xlab = "Age", ylab = "Death probability",
     pch = 19, frame = FALSE)

```
### Age and Gender vs Outcome
```{r}
NewTableAgeGender=dfNewTable %>% group_by(age_impute, gender) %>%
  summarise(expired=sum(hospitaldischargestatusId=="0"),total=sum(hospitaldischargestatusId %in%c("0","1")),percentage=expired/total*100) 
NewTableAgeGender<-na.omit(NewTableAgeGender)
NewTableAgeGender<-filter (NewTableAgeGender, gender %in% c("Male","Female"))
ggplot(NewTableAgeGender, aes(x=age_impute, y=percentage, group=gender)) +
  geom_point(aes(color=gender))


b <- c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
names=c("0-5","5-10","10-15","15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
NewTableAgeGender$BreakedAge=cut(NewTableAgeGender$age_impute, breaks = b, labels = names)
NewTableAgeGender<-na.omit(NewTableAgeGender)
SecondTable=NewTableAgeGender %>% group_by(BreakedAge,gender) %>%
  summarise(expiredTotal=sum(percentage)/5)
SecondTable<-na.omit(SecondTable)
ggplot(SecondTable, aes(x=BreakedAge, y=expiredTotal, group=gender)) +
  geom_line(aes(color=gender))+
  geom_point(aes(color=gender))


ggplot(SecondTable,aes(x=BreakedAge, y=expiredTotal,width=.45,fill=factor(gender)))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_discrete(name="Gender",
                      breaks=c(1, 2), 
                      labels=c("Male","Female")) +
  xlab("Gender")+ylab("Mortality Rate")
```


###disease-outcome
```{r}
###pop>50 and percentage>20
DiseaseTable=df %>% group_by(apacheadmissiondx) %>%
  summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100) 
DiseaseTable<-na.omit(DiseaseTable)
DiseaseTable<-filter (DiseaseTable, total>50)
DiseaseTable<-filter (DiseaseTable, percentage>20)


DiseaseTable$diseaseId=as.numeric(as.factor(DiseaseTable$apacheadmissiondx))

ggplot(data=DiseaseTable,aes(x=reorder(diseaseId,percentage),y=percentage)) + 
  geom_bar(stat ='identity',aes(fill=percentage))+
  coord_flip() + 
  theme_grey() + 
  scale_fill_gradient(name="percentage")+
  labs(title = 'disease vs percentage',
       y='percentage',x='disease')+ 
  geom_hline(yintercept = mean(DiseaseTable$percentage),size = 1, color = 'blue')

###pop >1000, percentage> 15

DiseaseTable=df %>% group_by(apacheadmissiondx) %>%
  summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100) 
DiseaseTable<-na.omit(DiseaseTable)
DiseaseTable<-filter (DiseaseTable, total>1000)
DiseaseTable<-filter (DiseaseTable, percentage>15)
DiseaseTable$diseaseId=as.numeric(as.factor(DiseaseTable$apacheadmissiondx))

ggplot(data=DiseaseTable,aes(x=reorder(diseaseId,percentage),y=percentage)) +
  geom_bar(stat ='identity',aes(fill=percentage))+
  coord_flip() +
  theme_grey() +
  scale_fill_gradient(name="percentage")+
  labs(title = 'disease vs percentage',
       y='percentage',x='disease')+
 geom_hline(yintercept = mean(DiseaseTable$percentage),size = 1, color = 'blue')
```




###Race-Outcome
```{r}
df_ethnicity =df %>%subset(ethnicity!='' )%>%group_by(ethnicity)%>%
   summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100)
ggplot(df_ethnicity, aes(x = ethnicity, y = percentage)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = ethnicity), vjust = -0.3) 
```
###Age-race-Outcome- 
```{r}
 ethnicityTable=df %>%subset(ethnicity!='')%>% group_by(ethnicity, age_impute)%>%
   summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100)
ethnicityTable<-filter (ethnicityTable, ethnicity %in% c("Caucasian","African American"))
ethnicityTable<-na.omit(ethnicityTable)
ggplot(ethnicityTable, aes(x=age_impute, y=percentage, group=ethnicity)) +
 geom_point(aes(color=ethnicity))#+
  #geom_line(aes(color=ethnicity))
 
# 
#######5years
b <- c(15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
names=c("15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
ethnicityTable$BreakedAge=cut(ethnicityTable$age_impute, breaks = b, labels = names)
ethnicityTable<-na.omit(ethnicityTable)

ethnicitySecondTable=ethnicityTable %>% group_by(BreakedAge,ethnicity) %>%
  summarise(Newexpired=sum(expired),Newtotal=sum(total),Newpercentage=Newexpired/Newtotal*100)
ggplot(ethnicitySecondTable, aes(x=BreakedAge, y=Newpercentage, group=ethnicity)) +
  geom_line(aes(color=ethnicity))+
  geom_point(aes(color=ethnicity))

ggplot(ethnicitySecondTable,aes(x=BreakedAge, y=Newpercentage,width=.45,fill=factor(ethnicity)))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_discrete(name="ethnicity",
                      breaks=c(1, 2), 
                      labels=c("Caucasian","African American")) +
  xlab("Age")+ylab("Mortality Rate")


```
```{r}
####10 years
b <- c(10,20,30,40,50,60,70,80,90)
names=c("10-20","20-30","30-40","40-50","50-60","60-70","70-80","80-90")
ethnicityTable$BreakedAge=cut(ethnicityTable$age_impute, breaks = b, labels = names)
ethnicityTable<-na.omit(ethnicityTable)

ethnicitySecondTable=ethnicityTable %>% group_by(BreakedAge,ethnicity) %>%
  summarise(Newexpired=sum(expired),Newtotal=sum(total),Newpercentage=Newexpired/Newtotal*100)
ggplot(ethnicitySecondTable, aes(x=BreakedAge, y=Newpercentage, group=ethnicity)) +
  geom_line(aes(color=ethnicity))+
  geom_point(aes(color=ethnicity))

ggplot(ethnicitySecondTable,aes(x=BreakedAge, y=Newpercentage,width=.45,fill=factor(ethnicity)))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_discrete(name="ethnicity") +
 #geom_text(aes(label = Newpercentage), size = 3, hjust = 0.5, vjust = 3)+
  xlab("Age")+ylab("Mortality Rate")

```

###F-TEST
```{r}
# dfEthOut=df%>%select(c(ethnicity,hospitaldischargestatus))
# dfEthOut <-na.omit(dfEthOut)
# dfEthOut=dfEthOut%>%mutate(output=ifelse(hospitaldischargestatus  %in% c("Expired"), 0, 1))
# White=dfEthOut%>%subset(ethnicity=="Caucasian")
# Black=dfEthOut%>%subset(ethnicity=="African American")
# 
# f.test(White$output,Black$output)
```
###checking relationship between age and race and disease
```{r}
########most common disease
##Sepsis, pulmonary
######## most fatal disease
##Cardiac arrest (with or without respiratory arrest; for respiratory arrest see Respiratory System)
df_ethnicityDiseas =df %>%subset(ethnicity %in% c("Caucasian","African American") )%>%subset(apacheadmissiondx=='Sepsis, pulmonary')%>%group_by(ethnicity)%>%
   summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100)
ggplot(df_ethnicityDiseas, aes(x = ethnicity, y = percentage)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = ethnicity), vjust = -0.3) 


ethnicityDiseaseTable=df %>%subset(ethnicity!='')%>%subset(apacheadmissiondx=='Sepsis, pulmonary')%>% group_by(ethnicity, age_impute)%>%
   summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100)
 ethnicityDiseaseTable<-filter (ethnicityDiseaseTable, ethnicity %in% c("Caucasian","African American"))
ethnicityDiseaseTable<-na.omit(ethnicityDiseaseTable)
ggplot(ethnicityDiseaseTable, aes(x=age_impute, y=percentage, group=ethnicity)) +
 geom_point(aes(color=ethnicity))


b <- c(15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
names=c("15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
ethnicityDiseaseTable$BreakedAge=cut(ethnicityDiseaseTable$age_impute, breaks = b, labels = names)
ethnicityDiseaseTable<-na.omit(ethnicityDiseaseTable)

ethnicityDiseaseSecondTable=ethnicityDiseaseTable %>% group_by(BreakedAge,ethnicity) %>%
  summarise(Newexpired=sum(expired),Newtotal=sum(total),Newpercentage=Newexpired/Newtotal*100)
ggplot(ethnicityDiseaseSecondTable, aes(x=BreakedAge, y=Newpercentage, group=ethnicity)) +
  geom_point(aes(color=ethnicity))

ggplot(ethnicityDiseaseSecondTable,aes(x=BreakedAge, y=Newpercentage,width=.45,fill=factor(ethnicity)))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_discrete(name="ethnicity",
                      breaks=c(1, 2), 
                      labels=c("Caucasian","African American")) +
  xlab("Race")+ylab("Mortality")
	
```
####Hospital vs output
```{r}

Hospital=df %>% group_by(hospitalid) %>%
  summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100) 
Hospital<-na.omit(Hospital)


##Hospital with more patients= 73
##Hospital with pation greator than 2000 and biggist death= 208
##Hospital with pation greator than 1000 and biggist death= 142
df_ethnicityHospital =df %>%subset(ethnicity %in% c("Caucasian","African American") )%>%subset(hospitalid=='73')%>%group_by(ethnicity)%>%
   summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100)
ggplot(df_ethnicityHospital, aes(x = ethnicity, y = percentage)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = ethnicity), vjust = -0.3) 


ethnicityHospitalTable=df %>%subset(ethnicity!='')%>%subset(hospitalid=='73')%>% group_by(ethnicity, age_impute)%>%
   summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100)
 ethnicityHospitalTable<-filter (ethnicityHospitalTable, ethnicity %in% c("Caucasian","African American"))
ethnicityHospitalTable<-na.omit(ethnicityHospitalTable)
ggplot(ethnicityHospitalTable, aes(x=age_impute, y=percentage, group=ethnicity)) +
 geom_point(aes(color=ethnicity))


b <- c(15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
names=c("15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
ethnicityHospitalTable$BreakedAge=cut(ethnicityHospitalTable$age_impute, breaks = b, labels = names)
ethnicityHospitalTable<-na.omit(ethnicityHospitalTable)

ethnicityHospitalSecondTable=ethnicityHospitalTable %>% group_by(BreakedAge,ethnicity) %>%
  summarise(Newexpired=sum(expired),Newtotal=sum(total),Newpercentage=Newexpired/Newtotal*100)
ggplot(ethnicityHospitalSecondTable, aes(x=BreakedAge, y=Newpercentage, group=ethnicity)) +
  geom_point(aes(color=ethnicity))


ggplot(ethnicityHospitalSecondTable,aes(x=BreakedAge, y=Newpercentage,width=.45,fill=factor(ethnicity)))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_discrete(name="ethnicity",
                      breaks=c(1, 2), 
                      labels=c("Caucasian","African American")) +
  xlab("Race")+ylab("Mortality")

```

