---
title: "Exploring Patient Table"
output: rmarkdown::github_document
---
Author: [Mahdi Moqri,Atefeh Rajabalizadeh]

The patinet table is a core part of the eICU-CRD and contains all information related to tracking patient unit stays. The table also contains patient demographics and hospital level information.


### Loading libraries
```{r message=FALSE}
library(data.table)
library(dplyr)
library(magrittr)
#library(plyr)
library(knitr)
library(ggplot2)
library(visreg)
library(rpart)
```

### Examining the patient table
The uniquePid column identifies a single patient across multiple stays. Let's look at a single uniquepid.

```{r}
# data_folder is the local path to your data
df=fread('patient.csv')
cols=c('uniquepid', 'gender', 'age', 'ethnicity','hospitaldischargestatus')
kable(head(df %>% select(cols)))
```
##patients
```{r}
#Unique patients
NumofPatient=length(unique(df$uniquepid))
NumofPatient
#unique hospital referral
NumofPatienttoHospital=length(unique(df$patienthealthsystemstayid))
NumofPatienttoHospital
#unique ICU referral
NumofPatienttoICU=length(unique(df$patientunitstayid))
NumofPatienttoICU
```

### Age
```{r message=FALSE, warning=FALSE}
df %<>% mutate(age_impute= (ifelse(age=='> 89',90, as.numeric(age))))
age=df[!is.na(df$age_impute),]$age_impute
hist(age)
median(age)
data.frame(table(df$age))
df_hist =df %>%subset(hospitaldischargestatus %in% c("Expired","Alive") )
# ggplot(df_hist, aes(x=age, fill=hospitaldischargestatus)) +
# geom_histogram(stat="count")     
b <- c(15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
names=c("15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
df_hist =df %>%subset(hospitaldischargestatus %in% c("Expired","Alive") )
df_hist$BreakedAge=cut(df_hist$age_impute, breaks = b, labels = names)
df_hist<-na.omit(df_hist)
SecondHistTable=df_hist %>% group_by(BreakedAge) 

ggplot(SecondHistTable, aes(x=BreakedAge, fill=hospitaldischargestatus)) +
geom_histogram(stat="count")  +
  xlab("Age")+ylab("Count of Patients")+
guides(fill=guide_legend(title="Discharge status"))+
scale_fill_manual("legend", values = c("Alive" = "#00AFBB", "Expired" = "#FF6666"))+theme(axis.text.x = element_text(angle = 90, hjust = 1, size=10,color="black"))
```

### Gender
```{r}
gender=table(df$gender)[c('Female','Male')]
gender
pie(gender)


SecondGenderTable=df%>%select (gender, hospitaldischargestatus)%>%subset(gender %in% c('Female','Male'))%>%subset(hospitaldischargestatus %in% c("Expired","Alive") ) %>% group_by(gender) 
SecondGenderTable=na.omit(SecondGenderTable)
ggplot(SecondGenderTable, aes(x=gender, fill=hospitaldischargestatus)) +
geom_histogram(stat="count")  +
  xlab("Gender")+ylab("Freq")

RealPatientGender=df%>%select(c(gender,uniquepid))%>%distinct(uniquepid, .keep_all = TRUE)
table(RealPatientGender$gender)

```

### Ethnicity
```{r, catche=TRUE}
df_ethnicity =df %>% subset(ethnicity!='')
ethnicity=df_ethnicity$ethnicity
kable(table(ethnicity))
barplot(table(ethnicity))
```
###Race-Outcome
```{r}
df_ethnicity =df %>%subset(ethnicity!='' )%>%group_by(ethnicity)%>%
   summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100)
ggplot(df_ethnicity, aes(x = ethnicity, y = percentage)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = ethnicity), vjust = -0.3) 
```
### Ethnicity
```{r, catche=TRUE}
df_ethnicity =df %>% subset(ethnicity!='')


df_ethnicity=as.data.frame(table(df_ethnicity$ethnicity))
df_ethnicity=df_ethnicity %>%
arrange(desc(Freq)) %>%
mutate(percentage = percent(Freq / sum(Freq))) 


ggplot(df_ethnicity, aes(x = "", y = Freq, fill = fct_inorder(Var1))) +
       geom_bar(width = 1, stat = "identity") +
       coord_polar("y", start = 0) +
  xlab("")+ylab("percentage")+
  theme(axis.text.x=element_blank()) + scale_fill_brewer(palette="Greens")+
       geom_label_repel(aes(label = percentage), size=5, show.legend = F, nudge_x = 1) +
       guides(fill = guide_legend(title = "Ethnicity"))
```
### Status at discharge
```{r}
unit_discharge=df$unitdischargestatus
table(unit_discharge)
pie(table(unit_discharge))
hostpial_discharge=df$hospitaldischargestatus
kable(table(hostpial_discharge))
pie(table(hostpial_discharge))
RealHospitalDeath=df%>%select(c(unitdischargestatus,patienthealthsystemstayid))%>%distinct(patienthealthsystemstayid, .keep_all = TRUE)
table(RealHospitalDeath$unitdischargestatus)
```



### Age vs Outcome
```{r message=FALSE, warning=FALSE}


df = df %>% mutate(hospitaldischargestatusId = ifelse(hospitaldischargestatus  %in% c("Expired"), 0, 1))
dfNewTable=df%>%select(c("hospitaldischargestatusId","gender","age_impute"))
NewTableAge=dfNewTable %>% group_by(age_impute) %>%
  summarise(expired=sum(hospitaldischargestatusId=="0"),total=sum(hospitaldischargestatusId %in%c("0","1")),percentage=expired/total*100) 
NewTableAge<-na.omit(NewTableAge)
plot(NewTableAge$`age_impute`, NewTableAge$percentage, main = "Age vs Death",
     xlab = "Age", ylab = "Death probability",
     pch = 19, frame = FALSE)

```
### Age and Gender vs Outcome
```{r message=FALSE, warning=FALSE}


df = df %>% mutate(hospitaldischargestatusId = ifelse(hospitaldischargestatus  %in% c("Expired"), 0, 1))
dfNewTable=df%>%select(c("hospitaldischargestatusId","gender","age_impute"))
NewTableAge=dfNewTable %>% group_by(age_impute) %>%
  summarise(expired=sum(hospitaldischargestatusId=="0"),total=sum(hospitaldischargestatusId %in%c("0","1")),percentage=expired/total*100) 
NewTableAge<-na.omit(NewTableAge)
NewTableAgeGender=dfNewTable %>% group_by(age_impute, gender) %>%
  summarise(expired=sum(hospitaldischargestatusId=="0"),total=sum(hospitaldischargestatusId %in%c("0","1")),percentage=expired/total*100) 
NewTableAgeGender<-na.omit(NewTableAgeGender)
NewTableAgeGender<-filter (NewTableAgeGender, gender %in% c("Male","Female"))
b <- c(15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
names=c("15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
NewTableAgeGender$BreakedAge=cut(NewTableAgeGender$age_impute, breaks = b, labels = names)
NewTableAgeGender<-na.omit(NewTableAgeGender)
SecondTable=NewTableAgeGender %>% group_by(BreakedAge,gender) %>%
  summarise(expiredTotal=sum(percentage)/5)
SecondTable<-na.omit(SecondTable)

ggplot(SecondTable,aes(x=BreakedAge, y=expiredTotal,width=.45,fill=factor(gender)))+
  geom_bar(stat="identity",position="dodge")+
  
  guides(fill=guide_legend(title="Gender"))+
scale_fill_manual("legend", values = c("Male" = "#00AFBB", "Female" = "#FF6666"))+
  xlab("Age")+ylab("Mortality Rate")+theme(axis.text.x = element_text(angle = 90, hjust = 1, size=10,color="black"))
```


###disease-outcome
```{r}
###pop>50 and percentage>20
DiseaseTable=df %>% group_by(apacheadmissiondx) %>%
  summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100) 
DiseaseTable<-na.omit(DiseaseTable)
DiseaseTable<-filter (DiseaseTable, total>50)
DiseaseTable<-filter (DiseaseTable, percentage>20)


DiseaseTable$diseaseId=as.numeric(as.factor(DiseaseTable$apacheadmissiondx))

ggplot(data=DiseaseTable,aes(x=reorder(diseaseId,percentage),y=percentage)) + 
  geom_bar(stat ='identity',aes(fill=percentage))+
  coord_flip() + 
  theme_grey() + 
  scale_fill_gradient(name="percentage")+
  labs(title = 'disease vs percentage',
       y='percentage',x='disease')+ 
  geom_hline(yintercept = mean(DiseaseTable$percentage),size = 1, color = 'blue')

###pop >2000, percentage> 15

DiseaseTable=df %>% group_by(apacheadmissiondx) %>%
  summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100) 
DiseaseTable<-na.omit(DiseaseTable)
DiseaseTable<-filter (DiseaseTable, total>4500)
DiseaseTable<-filter (DiseaseTable, percentage>10)
DiseaseTable$diseaseId=as.numeric(as.factor(DiseaseTable$apacheadmissiondx))

ggplot(data=DiseaseTable,aes(x=reorder(diseaseId,percentage),y=percentage)) +
  geom_bar(stat ='identity',aes(fill=percentage))+
  coord_flip() +
  theme_grey() +
  scale_fill_gradient(name="Mortality Rate")+
  labs(y='Mortality Rate',x='disease')+
 geom_hline(yintercept = mean(DiseaseTable$percentage),size = 1, color = 'blue')
```





###Age-race-Outcome- 
```{r}
 ethnicityTable=df %>%subset(ethnicity!='')%>% group_by(ethnicity, age_impute)%>%
   summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100)
ethnicityTable<-filter (ethnicityTable, ethnicity %in% c("Caucasian","African American"))
ethnicityTable<-na.omit(ethnicityTable)

b <- c(15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
names=c("15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
ethnicityTable$BreakedAge=cut(ethnicityTable$age_impute, breaks = b, labels = names)
ethnicityTable<-na.omit(ethnicityTable)

ethnicitySecondTable=ethnicityTable %>% group_by(BreakedAge,ethnicity) %>%
  summarise(Newexpired=sum(expired),Newtotal=sum(total),Newpercentage=Newexpired/Newtotal*100)


ggplot(ethnicitySecondTable,aes(x=BreakedAge, y=Newpercentage,width=.45,fill=factor(ethnicity)))+
  geom_bar(stat="identity",position="dodge")+
  
  guides(fill=guide_legend(title="ethnicity"))+
scale_fill_manual("legend", values = c("Caucasian" = "#00AFBB", "African American" = "#FF6666"))+
  xlab("Age")+ylab("Mortality Rate")+theme(axis.text.x = element_text(angle = 90, hjust = 1, size=10,color="black"))


```


