---
title: "Exploring admissionDrug Table"
output: rmarkdown::github_document
---
Author: [Mahdi Moqri](https://www.moqri.com/)

Adopted from https://github.com/mit-lcp/eicu-code

The following columns are available:

* admissiondrugid - primary key, has no meaning but identifies rows uniquely
* drugOffset - number of minutes from unit admit time that the admission drug was administered
* drugEnteredOffset - number of minutes from unit admit time that the admission drug was entered
* drugNoteType - unique note picklist types e.g.: Comprehensive Progress Admission Intubation
* specialtyType - physician specialty picklist types e.g.: anesthesiology gastroenterology oncology
* userType - who documented the drug from eCareManager user picklist types e.g.: eICU Physician, Nurse, Attending Physician
* rxincluded - Does the Note have associated Rx data: True or False
* writtenIneICU - Was the Note written in the eICU: True or False
* drugName - name of the selected admission drug e.g.: POTASSIUM CHLORIDE/D5NS METAXALONE PRAVACHOL
* drugDosage - dosage of the admission drug e.g.: 20.0000 400.000
* drugUnit - picklist units of the admission drug e.g.: mg mg/kg patch
* drugAdmitFrequency - picklist frequency with which the admission drug is administred e.g.: PRN twice a day at bedtime
* drughiclseqno - a code representing the drug (hierarchical ingredient code list, HICL)

We recommend configuring the `config.ini` file to allow for connection to the database without specifying your password each time.

```{r include=FALSE}
data_folder='/Users/mahdimoqri/data/eicu/csv'
```

### Loading libraries
```{r message=FALSE}
library(data.table)
library(dplyr)
library(magrittr)
library(plyr)
library(knitr)
```

### Examining the table
```{r}
# data_folder is the local path to your data
df=fread(file.path(data_folder,'admissiondrug.csv'))
kable(head(df))
```


### Examine a single patient¶
```{r}
patientunitstayid_ = 2704494
df_patientunitstayid=df %>% subset (patientunitstayid==patientunitstayid_)
head(df_patientunitstayid)
cols = c('admissiondrugid','patientunitstayid','drugoffset','drugenteredoffset','drugname','drughiclseqno')
df_patientunitstayid %>% select (cols)
```

Here we can see that these drugs were documented 2153 minutes (1.5 days) after ICU admission, but administered 87132 minutes (60 days) *before* ICU admission (thus, the negative offset). Since it's reasonable to assume the patient is still taking the drug (as this is the admissiondrug table), `drugoffset` can likely be treated as a start time for a prescription of the drug.

### Identifying patients admitted on a single drug
Let's look for patients who were admitted on Zaroxolyn.

```{r}
drug = 'ZAROXOLYN'
cols=c('admissiondrugid', 'patientunitstayid'
  , 'drugoffset', 'drugenteredoffset'
  , 'drugname', 'drughiclseqno')
df_drug_name = df %>% subset (drugname==drug) %>% select (cols)
cat(length(unique(df_drug_name$patientunitstayid)),'unit stays with', drug)
```

Instead of using the drug name, we could try to use the HICL code.

```{r}
hicl = 3663
df_drug_hicl = df %>% subset (drughiclseqno==hicl) %>% select (cols)
cat(length(unique(df_drug_hicl$patientunitstayid)),'unit stays with HICL =', hicl)
```


As we can see, using the HICL returned many more observations. Let's take a look at a few:

```{r}
# rows in HICL which are *not* in the drug dataframe
idx = setdiff(df_drug_hicl$admissiondrugid,df_drug_name$admissiondrugid)
# count the drug names
table(df_drug_hicl %>% subset (admissiondrugid %in% idx) %>% select(drugname))
```

All the rows use the drug name "Metolazone". Metolazone is the generic name for the brand Zaroxolyn. This demonstrates the utility of using HICL codes to identify drugs - synonyms like these are very common and can be tedious to find.

### Hospitals with data available¶
```{r}
patient=fread(file.path(data_folder,'patient.csv'))


patient_admission=patient %>% left_join(df) 
patient_group=patient_admission %>% group_by(hospitalid) %>% tally(name = "patients")
df_group=patient_admission %>% subset(!is.na(admissiondrugid)) %>% group_by(hospitalid) %>% tally(name = "drugs")

hospitals= patient_group %>% full_join(df_group)
hospitals=hospitals[order(-hospitals$drugs),]
hospitals=hospitals %>% mutate (data_completion=drugs/patients*100)
head(hospitals)
hospitals[is.na(hospitals)] = 0
hist(hospitals$data_completion,xlab = 'Number of hospitals', ylab = 'Percent of patients with data')
```

As is common in eICU-CRD, there are a subset of hospitals who routinely utilize this portion of the medical record (and thus have 90-100% data completion), while there are other hospitals who rarely use this interface and thus have poor data completion (0-10%).

