---
title: "Exploring Patient Table"
output: rmarkdown::github_document
---
Author: [Mahdi Moqri](https://www.moqri.com/)

Adopted from https://github.com/mit-lcp/eicu-code

The patinet table is a core part of the eICU-CRD and contains all information related to tracking patient unit stays. The table also contains patient demographics and hospital level information.

```{r include=FALSE}
data_folder='/Users/mahdimoqri/data/eicu/csv'
```

### Loading libraries
```{r message=FALSE}
library(data.table)
library(dplyr)
library(magrittr)
library(plyr)
library(knitr)
```

### Examining the patient table
The uniquePid column identifies a single patient across multiple stays. Let's look at a single uniquepid.

```{r}
# data_folder is the local path to your data
df=fread(file.path(data_folder,'patient.csv'))
kable(head(df))
```


### Examine a single patient¶
```{r}
uniquepid_ = '002-33870'
kable(df %>% subset(uniquepid==uniquepid_))
```

Here we see two unit stays for a single patient. Note also that both unit stays have the same patienthealthsystemstayid - this indicates that they occurred within the same hospitalization.

We can see the unitstaytype was 'admit' for one stay, and 'stepdown/other' for another. Other columns can give us more information.

### Identifying patients admitted on a single drug
Let's look for patients who were admitted on Zaroxolyn.

```{r}
cols=c('patientunitstayid', 'wardid', 'unittype', 'unitstaytype', 'hospitaladmitoffset', 'unitdischargeoffset')
kable(df %>% subset(uniquepid==uniquepid_) %>% select(cols) )
```

Note that it's not explicitly obvious which stay occurred first. Earlier stays will be closer to hospital admission, and therefore have a higher hospitaladmitoffset. Above, the stay with a hospitaladmitoffset of -14 was first (occurring 14 minutes after hospital admission), followed by the next stay with a hospitaladmitoffset of -22 (which occurred 22 minutes after hospital admission). Practically, we wouldn't consider the first admission a "real" ICU stay, and it's likely an idiosyncrasy of the administration system at this particular hospital. Notice how both rows have the same wardid.

### Patients' ages

As ages over 89 are required to be deidentified by HIPAA, the age column is actually a string field, with ages over 89 replaced with the string value '> 89'.

```{r}
kable(head(df %>% group_by(age) %>% tally() %>% arrange(-n)))
```


As is common in eICU-CRD, there are a subset of hospitals who routinely utilize this portion of the medical record (and thus have 90-100% data completion), while there are other hospitals who rarely use this interface and thus have poor data completion (0-10%).