---
title: "03-Racial Bias"
output: html_document
---


# Racial Bias

In this chapter, we first compare the mortality rate in different categories of ethnicity to make sure that all people receive the same care services.

## Ethnicity-Mortality

By comparing the mortality rate between different ethnicity categories, we can see that native American has the lowest mortality rate and Hispanic people have the highest. Among the 2 biggest categories, it is shown that they caucasian people have a slightly higher mortality rate in comparison to African American people. We will investigate this further to see at each age category they are different. 

```{r ethnicitycomp, message=FALSE, warning=FALSE, echo=FALSE}
df_ethnicity =df %>%subset(ethnicity!='' )%>%group_by(ethnicity)%>%
   summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100)
ggplot(df_ethnicity, aes(x = ethnicity, y = percentage)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  xlab("Ethnicity")+ylab("Mortality Rate")
```


## Ethnicity-Mortality-Age
As we saw in chapter 2, Caucasians and African American are the two biggest groups in patients' data. 
Figure below compares morality rate in black people versus white people in different age categories. Although the Caucasians people have slightely higher mortality rate than African Americans, we can see that at each age category, the rate of morality in black people is higher than white people.

```{r ethnicity age,message=FALSE, warning=FALSE, echo=FALSE}
 ethnicityTable=df %>%subset(ethnicity!='')%>% group_by(ethnicity, age_impute)%>%
   summarise(expired=sum(hospitaldischargestatus=="Expired"),total=sum(hospitaldischargestatus %in%c("Expired","Alive")),percentage=expired/total*100)
ethnicityTable<-filter (ethnicityTable, ethnicity %in% c("Caucasian","African American"))
ethnicityTable<-na.omit(ethnicityTable)

b <- c(15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90)
names=c("15-20","20-25","25-30","30-35","35-40","40-45","45-50","50-55","55-60","60-65","65-70","70-75","75-80","80-85","85-90")
ethnicityTable$BreakedAge=cut(ethnicityTable$age_impute, breaks = b, labels = names)
ethnicityTable<-na.omit(ethnicityTable)

ethnicitySecondTable=ethnicityTable %>% group_by(BreakedAge,ethnicity) %>%
  summarise(Newexpired=sum(expired),Newtotal=sum(total),Newpercentage=Newexpired/Newtotal*100)


ggplot(ethnicitySecondTable,aes(x=BreakedAge, y=Newpercentage,width=.45,fill=factor(ethnicity)))+
  geom_bar(stat="identity",position="dodge")+
  
  guides(fill=guide_legend(title="ethnicity"))+
scale_fill_manual("legend", values = c("Caucasian" = "#00AFBB", "African American" = "#FF6666"))+
  xlab("Age")+ylab("Mortality Rate")+theme(axis.text.x = element_text(angle = 90, hjust = 1, size=10,color="black"))


```






